Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-03-27 manuscript.Rmd
abstract

abstract: |
  Abstract

keywords          : "body position, motor development, everyday experiences, sitting, machine learning"
wordcount         : "X"

>>>>

abstract: |
  Long form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored. How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position? The aim of the current study was to validate long form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and hips. Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior. Moreover, classification remained accurate when predicting behavior later in the session when infants and caregivers were unsupervised and went about their normal activities, showing that the method can handle the challenge of measuring unconstrained, natural activity. Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors. We end the paper with a discussion of the novel opportunities that long form motor recordings afford for understanding infant learning and development.

keywords          : "body position, motor development, everyday experiences, sitting, machine learning"
wordcount         : "12,030"

### 2023-03-27 manuscript.tex
abstract

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\raggedbottom
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Long-form recording of infant body position in the home using wearable inertial sensors},
  pdfauthor={John M. Franchak1, Maximilian Tang1, Hailey Rousey1, \& Chuan Luo1},
  pdflang={en-EN},
  pdfkeywords={body position, motor development, everyday experiences, sitting, machine learning},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\abstract{%
Abstract
}

@@

What underpins these novel research applications is that body position classifications can be applied automatically at scale. Annotating a video corpus of infant behavior from a moderately sized sample---such as 40 infants recorded for 2 hours each in 2 separate visits (Herzberg et al., 2021)---is incredibly labor intensive. Annotating a larger, more representative sample of \emph{hundreds of infants} while simultaneously scoring \emph{full-day data} would be prohibitive. With sufficiently accurate group models, researchers could annotate a moderate video sample to then apply the model to hundreds of full-day recordings. The added cost of more recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability. Prior diary methods show that infants inconsistently display new motor skills on a day-to-day basis (Adolph et al., 2008; Adolph \& Robinson, 2011). Multi-day recordings could further uncover how infants' experiences vary over timescales from seconds to days.

@@

We acknowledge several limitations in the current approach. First, a stronger test of full-day accuracy would be to compare accuracy at the start of the recording period to accuracy at the end, rather than using the distal comparison video that followed the first 90 minutes of the study. As we described in the procedure section, synchronizing video cameras and inertial sensors required capturing a synchronization point on video. We completed this procedure prior to giving equipment to caregivers so that they were not responsible for synchronization; in fact, caregivers never touched a button on the video cameras. An end-of-day video would require a lot on the part of caregivers, and any mistakes in the procedure would lead to misaligned and unusable data. Given that most recordings ended at infants' bedtimes, having an experimenter visit families in the evening to video record would be intrusive.

@@

In summary, the current study demonstrates the validity of long-form recordings of infant motor behavior in the home. Our analyses show that body position classifications---whether infants are supine, prone, sitting, upright, or held---are accurate immediately and even following a substantial delay. Most important, we find substantial agreement between human-coded and model-predicted body position for data collected during truly everyday, unsupervised activities that create the most challenging cases for automatic classification. In most cases, model prediction accuracy approached human reliability, suggesting that model predictions can be confidently used in analyses of full-day infant behavior. Examining the resulting corpus of 206.66 hours of real-time infant body position showed the feasibility of capturing data covering the majority of infants' awake time and the variety of activities contained therein. A rudimentary analysis of aggregated data showed that full-day position estimates conformed to expectations about age differences in motor behavior. Future work employing the method can go beyond aggregated measures of behavior to uncover the temporal structure in infants' daily motor experiences.

@@

\leavevmode\vadjust pre{\hypertarget{ref-CasillasElliott2021}{}}%
Casillas, M., \& Elliott, M. (2021). \emph{Cross-cultural differences in children's object handling at home}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-DanchoVaughan2023}{}}%
Dancho, M., \& Vaughan, D. (2023). \emph{Timetk: A tool kit for working with time series in r}.

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: 12,030}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\raggedbottom
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Long-form recording of infant body position in the home using wearable inertial sensors},
  pdfauthor={John M. Franchak1, Maximilian Tang1, Hailey Rousey1, \& Chuan Luo1},
  pdflang={en-EN},
  pdfkeywords={body position, motor development, everyday experiences, sitting, machine learning},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\abstract{%
Long form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored. How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position? The aim of the current study was to validate long form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and hips. Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior. Moreover, classification remained accurate when predicting behavior later in the session when infants and caregivers were unsupervised and went about their normal activities, showing that the method can handle the challenge of measuring unconstrained, natural activity. Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors. We end the paper with a discussion of the novel opportunities that long form motor recordings afford for understanding infant learning and development.
}

@@

What underpins these novel research applications is that body position classifications can be applied automatically at scale. Annotating a video corpus of infant behavior from a moderately sized sample---such as 40 infants recorded for 2 hours each in 2 separate visits (Herzberg et al., 2021)---is incredibly labor intensive. Annotating a larger, more representative sample of \emph{hundreds of infants} while simultaneously scoring \emph{full-day data} would be prohibitive. With sufficiently accurate group models, researchers could annotate a moderately-sized video corpus to then apply the model to hundreds of full-day recordings. Unlike with video, the added cost of more IMU sensor recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability. Prior diary methods show that infants inconsistently display new motor skills on a day-to-day basis (Adolph et al., 2008; Adolph \& Robinson, 2011). Multi-day recordings could further uncover how infants' experiences vary over timescales from seconds to days.

@@

We acknowledge several limitations in the current approach. First, a stronger test of full-day accuracy would be to compare accuracy at the start of the recording period to accuracy at the end, rather than using the distal comparison video that followed the first 90 minutes of the study. As we described in the procedure section, synchronizing video cameras and inertial sensors required capturing a synchronization point on video. We completed this procedure prior to giving equipment to caregivers so that they were not responsible for synchronization; in fact, caregivers never (purposefully) touched a button on the video camera. An end-of-day video would require a lot of effort and compliance on the part of caregivers, and any mistakes in the procedure would lead to misaligned and unusable data. Given that most recordings ended at infants' bedtimes, having an experimenter visit families in the evening to video record would be intrusive.

@@

In summary, the current study demonstrates the validity of long-form recordings of infant motor behavior in the home. Our analyses show that body position classifications---whether infants are supine, prone, sitting, upright, or held---are accurate immediately and even following a substantial delay. Most important, we find substantial agreement between human-coded and model-predicted body position for data collected during truly everyday, unsupervised activities that create the most challenging cases for automatic classification. In most cases, model prediction accuracy approached human reliability, suggesting that model predictions can be confidently used in analyses of full-day infant behavior. Examining the resulting corpus of \textgreater{} 200 hours of real-time infant body position showed the feasibility of capturing data covering the majority of infants' awake time and the variety of activities contained therein. A rudimentary analysis of aggregated data showed that full-day position estimates conformed to expectations about age differences in motor behavior. Future work employing the method can go beyond aggregated measures of behavior to uncover the temporal structure in infants' daily motor experiences.

@@

\leavevmode\vadjust pre{\hypertarget{ref-CasillasElliott2021}{}}%
Casillas, M., \& Elliott, M. (2021). Cross-cultural differences in children's object handling at home. \emph{PsyArXiv}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-DanchoVaughan2023}{}}%
Dancho, M., \& Vaughan, D. (2023). \emph{Timetk: A tool kit for working with time series in {R}}.
